---
title: "Passing Rates"
author: "Aditya Kothari"
date: "June 2019"
output: md_document
---

```{r Setup, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'hide'}
   
rm(list = ls())

library(data.table)
library(emdist)
library(snow)
library(snowfall)
library(ggplot2)
library(scales)
library(knitr)
library(clue)
library(dendextend)
theme_set(theme_bw(12))

knitr::knit_meta(class=NULL, clean = TRUE)
options(knitr.duplicate.label = 'allow')

```

```{r ParametersFunctions, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'hide'}

cPostCodeFolderLocation = '/media/ask/Data/Personal/Projects/Personal/AnalysingFootballData/EMDRelated/'

source(paste0(cPostCodeFolderLocation, '/Common.R'))
read_chunk(paste0(cPostCodeFolderLocation, '/CommonChunks.R'))
read_chunk(paste0(cPostCodeFolderLocation, '/CommonChunksClusteringPlayersRoles.R'))

vnQuantiles = c(0.1, 0.5, 0.9)

```

```{r LoadingProcessedData, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'asis'}
```

```{r ProcessingMore, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'asis'}

dtPassingSpeed = dtPasses[,
   list(
      PassCountInPlay = .N, 
      Duration = diff(range((expandedMinute * 60) + second))
   ),
   list(
      Tournament,
      teamId,
      matchId,
      playId
   )
]

dtPassingSpeed = merge(
   dtPassingSpeed,
   dtMatches[, list(teamName = teamName[1]), teamId],
   'teamId'
)

```

Someone on the internet was repeating the EPL being very fast and players might struggle to cope with it, etc. spiel. I've been playing with some passing data over <a href = "https://thecomeonman.github.io/SpatialSimilaritiesBetweenPlayers/index.html">here</a> and thought I'd take a quick look at whether the passing reflects this statement or not.

( Unfortunately, the resolution to time in my dataset is only a second so some precision is lost. )

```{r TournamentCountMedians, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'asis'}

ggplot(dtPassingSpeed[Duration > 0][, list(PlayCount = .N, Low = quantile(PassCountInPlay / Duration, vnQuantiles[1]), Mid = quantile(PassCountInPlay / Duration, vnQuantiles[2]), High = quantile(PassCountInPlay / Duration, vnQuantiles[3])), list(Tournament, PassCountInPlay)][PlayCount > 50]) +
    geom_ribbon(aes(x = PassCountInPlay, ymax = High, ymin = Low, group = paste(Tournament), fill = Tournament), alpha = 0.25)  +
    geom_line(aes(x = PassCountInPlay, y = Mid,  group = paste(Tournament), color = Tournament)) +
    geom_line(aes(x = PassCountInPlay, y = High,  group = paste(Tournament), color = Tournament)) +
    geom_line(aes(x = PassCountInPlay, y = Low,  group = paste(Tournament), color = Tournament)) +
    geom_point(aes(x = PassCountInPlay, y = Mid, group = paste(Tournament), size = PlayCount)) + 
    labs(
        caption = 'Area boundaries are the 10th and 90th percentile of the respective value\nPrepared by thecomeonman',
        y = 'Passes per second',
        x = 'Number of passes in the play'
    )


```

Across leagues, the passing rate tends to settle at nearly the same value of a little lesser than 0.4 passes per second. The EPL doesn't look very different from the other leagues.

Since the rate tends to settle after a play of 5 passes or longer, I've only included plays of 5 passes or longer in the below charts. I also chopped off the data after plays of 13 passes because data was very scarce there. The points plotted satisfy some minimum criterion of number of plays with that many passes because otherwise the median would be very skewed.

When we split the data further into teams -

```{r TeamCountMedians, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 15, results = 'asis'}

ggplot(
   dtPassingSpeed[
      Duration > 0
   ][, 
      list(
         PlayCount = .N, 
         Low = quantile(PassCountInPlay / Duration, vnQuantiles[1]), 
         Mid = quantile(PassCountInPlay / Duration, vnQuantiles[2]),
         High = quantile(PassCountInPlay / Duration, vnQuantiles[3])
      ), 
      list(Tournament, PassCountInPlay, teamId, teamName)
   ][
      PlayCount > 10
   ][
      PassCountInPlay > 4 &
      PassCountInPlay < 14
   ]
) +
   # geom_ribbon(aes(x = PassCountInPlay, ymax = High, ymin = Low, group = paste(Tournament), fill = Tournament), alpha = 0.25)  +
   geom_jitter(aes(x = Tournament, y = Mid,  group = paste(Tournament), color = Tournament, size = PlayCount), height = 0, width = 0.25) +
   facet_wrap(~PassCountInPlay, labeller = label_both) +
   labs(
      caption = 'Prepared by thecomeonman',
      y = 'Median Passes per second',
      x = 'Tournament',
      title = 'Passing speed, data from 2017/18'
   )
```

Each point is the median passing rate for the respective team. Each box is the number of passes in the play.

Based on this chart, the EPL might be sliiiiightly faster than the other 5 leagues. The Bundesliga clearly looks the slowest though.

For the fun of it, I'm going to dig a little deeper. Adding labels to the previous charts to be able to discern the teams -

```{r TeamCountMedianWithLabels, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 15, results = 'asis'}

ggplot(
   dtPassingSpeed[
      Duration > 0
   ][, 
      list(
         PlayCount = .N, 
         Low = quantile(PassCountInPlay / Duration, vnQuantiles[1]), 
         Mid = quantile(PassCountInPlay / Duration, vnQuantiles[2]),
         High = quantile(PassCountInPlay / Duration, vnQuantiles[3])
      ), 
      list(Tournament, PassCountInPlay, teamId, teamName)
   ][
      PlayCount > 10
   ][
      PassCountInPlay > 4 &
      PassCountInPlay < 14
   ]
) +
   # geom_ribbon(aes(x = PassCountInPlay, ymax = High, ymin = Low, group = paste(Tournament), fill = Tournament), alpha = 0.25)  +
   geom_jitter(aes(x = Tournament, y = Mid,  group = paste(Tournament), color = Tournament, size = PlayCount), height = 0.01, width = 0.25) +
   geom_text(aes(x = Tournament, y = Mid,  group = paste(Tournament), label = teamName), size = 3, position = position_jitter(height = 0.02, width = 0.25)) +
   facet_wrap(~PassCountInPlay, labeller = label_both) +
   labs(
      caption = 'Prepared by thecomeonman',
      y = 'Median passes per second',
      x = 'Tournament',
      title = 'Passing speed, data from 2017/18'
   )

```

It's a little too cluttered. I'll combine all the plays of differing passing lengths, to make it easier to consume -

```{r TeamMedianWithLabels, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'asis'}

ggplot(
   dtPassingSpeed[
      PassCountInPlay > 4
   ][
      Duration > 0
   ][, 
      list(
         PlayCount = .N, 
         Low = quantile(PassCountInPlay / Duration, vnQuantiles[1]), 
         Mid = quantile(PassCountInPlay / Duration, vnQuantiles[2]),
         High = quantile(PassCountInPlay / Duration, vnQuantiles[3])
      ), 
      list(Tournament, teamId, teamName)
   ]
) +
   # geom_ribbon(aes(x = PassCountInPlay, ymax = High, ymin = Low, group = paste(Tournament), fill = Tournament), alpha = 0.25)  +
   geom_jitter(aes(x = Tournament, y = Mid,  group = paste(Tournament), color = Tournament, size = PlayCount), height = 0.00, width = 0.05) +
   geom_text(aes(x = Tournament, y = Mid,  group = paste(Tournament), label = teamName), size = 4, position = position_jitter(height = 0.003, width = 0.25)) +
   labs(
      caption = 'Prepared by thecomeonman',
      y = 'Median passes per second',
      x = 'Tournament',
      title = 'Passing speed, data from 2017/18'
   )

```

Sarri's Napoli heads and shoulders above everybody else. Reading looking the worst of them all.

The difference between Napoli and the overall median is probably lesser than one extra pass every ten seconds though. That's all it takes for us to perceive Sarriball as high tempo football.


```{r Discard1, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 15, results = 'asis'}

if ( F ) {
   
   ggplot(
      dtPassingSpeed[
         PassCountInPlay > 4
      ][
         Duration > 0
      ][, 
         list(
            PlayCount = .N, 
            Low = quantile(PassCountInPlay / Duration, vnQuantiles[1]), 
            Mid = quantile(PassCountInPlay / Duration, vnQuantiles[2]),
            High = quantile(PassCountInPlay / Duration, vnQuantiles[3])
         ), 
         list(Tournament, teamId, teamName)
      ]
   ) +
      # geom_ribbon(aes(x = PassCountInPlay, ymax = High, ymin = Low, group = paste(Tournament), fill = Tournament), alpha = 0.25)  +
      geom_point(aes(x = teamName, y = Mid,  group = paste(Tournament), color = Tournament, size = PlayCount)) +
      geom_point(aes(x = teamName, y = Low,  group = paste(Tournament), color = Tournament)) +
      geom_point(aes(x = teamName, y = High,  group = paste(Tournament), color = Tournament)) +
      labs(
         caption = 'Prepared by thecomeonman',
         y = 'Median passes per second',
         x = 'Tournament',
         title = 'Passing speed, data from 2017/18'
      ) + 
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0))

}

```





```{r Discard2, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 15, results = 'asis'}

if ( F ) {

   ggplot(
      merge(
         dtPassingSpeed[
              Duration > 0
         ][, 
              list(
                 PlayCount = .N, 
                 Low = quantile(PassCountInPlay / Duration, vnQuantiles[1]), 
                 Mid = quantile(PassCountInPlay / Duration, vnQuantiles[2]), 
                 High = quantile(PassCountInPlay / Duration, vnQuantiles[3])
              ), 
              list(
                  Tournament, teamId, PassCountInPlay
              )
          ][
              PlayCount > 10
          ][
              PassCountInPlay > 5 & PassCountInPlay < 12
          ][,
             .SD[order(Mid)],
             PassCountInPlay
          ][,
               Order := .I - min(.I) + 1,
               PassCountInPlay
         ],
         dtMatches[, list(teamName = teamName[1]), teamId],
         'teamId'
       )
   ) +
      geom_line(
          aes(
             # x = Tournament, 
             x = PassCountInPlay,
             y = Mid, 
             group = paste(Tournament, teamId), 
             # color = Tournament, 
             # size = PlayCount
            ), 
          width = 0.25, 
          height = 0.01
      ) + 
      # geom_text(
      #     aes(
      #        x = teamName, 
      #        # x = Tournament, 
      #        y = Mid, 
      #        label = teamName
      #       ),
      #     position = position_jitter(
      #        width = 0.25, height = 0.01
      #     )
      # ) + 
       labs(
           caption = ' - thecomeonman',
           y = 'Median passes per second',
           x = 'Number of passes in the play'
       ) + 
       # facet_wrap(~PassCountInPlay, ncol =  3) + 
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0))

}
```